{\acronym}: a machine learning based approach to entity linking combining different features

### 2015-08-21 paper.tex
- various small fixes

%opening
\title{{\acronym} meets FOX: a combined approach to entity linking based on vector similarity }
\titlerunning{{\acronym} meets FOX}

@@

The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{{\acronym}}, a novel approach that is based on a combination of different features that is trained using a Support Vector Machine.  Given a named entity and a resource candidate, {\acronym} combines four features that quantify i) the likelihood that the actual name refers to the resource candidate (extracted from Wikipedia anchors), ii) the similarity of the textual context in which the name occurs to the textual context of the resource candidate (measured as the bag-of-words similarity using cosine), iii) the overlap between the semantic neighborhood signature of the resource candidate and the named entities mentioned in the larger context of the named entity, and iv) the prominence of the resource candidate as measured by its page rank in the DBpedia graph. An optimal non-linear model is trained using an SVM which combines these four features into an overall decision whether the named entity refers to a candidate resource or not.

@@

The input data for {\acronym} is a regular piece of text, for example a newspaper article, in which entity-representing sequences of tokens (\emph{fragments}) are already identified (we use the Stanford NER~\cite{ner}). For a given fragment, the algorithm's task is to determine the matching real-world entity. It does so by picking the most likely entity from a set of entity candidates, e.g. specified in a knowledge base, in our case DBpedia.

@@

Candidates are scored by a support vector machine.\footnote{{\acronym} is using Weka's SMO: \url{http://weka.sourceforge.net/doc.dev/weka/classifiers/functions/SMO.html}} Four features are computed for each candidate, which are used as input for the SVM: tf-idf vector similarity, semantic neighborhood signature vector similarity, candidate reference probability, and page rank. In the following we explain each of them in more detail.

% ------------ tf--idf vector similarity -------------------
\subsection{tf-idf Vector Similarity}

@@

To calculate the similarity score for each candidate, their semantic neighborhood signature vector needs to compared to the semantic neighborhood signature vector of the input. Unfortunately, the input does not have a semantic neighborhood signature of its own. As a surrogate a candidate vector is generated as follows: Let $m(e) \rightarrow c$ be a mapping from an entity $e$ to a count $c$. For each candidate entity of every fragment, check whether it already appears in the mapping. If not, add it to the map and set its count to one. If it exists, increment its count by one.

@@

$$\mathrm{crp}(a, e) := \frac{\mathrm{ref}(a, e)}{{\sum\limits_{a'\in E} \mathrm{ref}(a', e)}}$$

Where $\mathrm{ref}(a, e)$ is the number of references from anchor/candidate $a$ to entity $e$ and $E$ is the set of all anchors referring to $e$.

@@

The idea behind the PageRank~\cite{pageRank} algorithm is to estimate the relevance of a website to a search query not only by whether the search terms appear on that page, but also by how many other pages were providing a hyperlink to the website in question. As DBpedia entries form a highly interconnected graph, the same principle can be applied in order to determine the page rank of a resource in DBpedia.

Let $u$ be a web page. $B_u$ is the set of pages that have a hyperlink pointing to $u$. Let $N_u$ be the number of outgoing links for page $u$. A slightly simplified PageRank is then defined as follows: 
$$R(u) = c \sum_{v \in B_u} \frac{R(v)}{N_v}$$
Where $c$ is a normalizing factor. Notably the PageRank of a page is derived from the PageRank of its incoming links, meaning that the algorithm has to be iterated until the system reaches a steady state. The overall PageRank for a graph is supposed to be constant, but there are corner cases (like cycles) that have to be accounted for; see \cite{pageRank} for more information. Our implementation uses the JUNG framework\footnote{Java Universal Network/Graph Framework: \url{http://jung.sourceforge.net/}}.

@@

Table~\ref{table:redirects1} shows the evaluation results. 
\acronym{} outperforms both FOX and DBpedia Spotlight on the AIDA/CoNLL dataset and achieves slightly worse but comparable results on the KORE50 dataset. 
The latter is due to the fact that KORE50 was designed to have a high degree of ambiguity and short sentences, which renders {\acronym 's} approach less effective: Short sentences provide fewer tokens to evaluate a tf-idf overlap, while increased numbers of candidates allow for more detrimental semantic neighborhood signature vector overlaps.

@@

@@

\begin{table}[t]
\begin{center}
\begin{tabularx}{\textwidth}{ X Y Y Y Y Y Y}
\toprule
& \multicolumn{3}{c}{AIDA/CoNLL}  
& \multicolumn{3}{c}{KORE50}\\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
 & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ 
\midrule
\emph{without} &  &  &  &  &  &  \\ 
SNS & 0.60 & 0.57 & 0.59 & 0.32 & 0.28 &  0.29 \\
tf-idf & 0.60 & 0.57 & 0.58 & 0.32 & 0.28 &  0.29 \\
page rank & 0.36 & 0.34 & .35 & 0.12 & 0.10 &  0.11 \\
CRP & 0.62 & 0.60 & 0.61 & 0.31 & 0.27 & 0.29 \\
\midrule
\emph{only with} &  &  &  &  &  &  \\ 
SNS & 0.36 & 0.35 & 0.36 & 0.07 & 0.06 &  0.06 \\
tf-idf & 0.22 & 0.21 & 0.22 & 0.08 & 0.07 &  0.07 \\
page rank & 0.62 & 0.60 & 0.61 & 0.32 & 0.28 &  0.29 \\
CRP & 0.21 & 0.20 & 0.21 & 0.08 & 0.07 &  0.07 \\
\bottomrule
\end{tabularx}
\end{center}
\caption{Results for the AIDA/CoNLL and KORE50 datasets when one of the components is removed and when only one component is used for classification. SNS is Semantic Neighborhood Signature Vector Similarity, CRP is Candidate Reference Probability.}
\label{table:components}
\end{table}

@@

We have presented {\acronym}, a new entity linking approach that exploits a combination of features that are linearly combined using a Support Vector Machine. We have shown that the approach works very well when the contexts are longer, clearly outperforming FOX and DBpedia Spotlight on the AIDA CoNLL dataset. On the KORE50 dataset where contexts are shorter, it achieves comparable results to these approaches. In future work, we will release a service that can be publicly accessed using a similar interface to DBpedia Spotlight and FOX. We will also extend the approach to other languages and perform a more systematic evaluation using the GERBIL framework \cite{gerbil}.

>>>>

%opening
\title{{\acronym}: a machine learning based approach to entity linking combining different features}
\titlerunning{{\acronym}}

@@

The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{{\acronym}}, a novel approach that is based on a combination of different features that is trained using a Support Vector Machine.  Given a named entity and a resource candidate, {\acronym} combines four features that quantify i) the likelihood that the actual name refers to the resource candidate (extracted from Wikipedia anchors), ii) the similarity of the textual context in which the name occurs to the textual context of the resource candidate (measured as the bag-of-words similarity using cosine), iii) the overlap between the semantic neighborhood signature of the resource candidate and the named entities mentioned in the larger context of the named entity, and iv) the prominence of the resource candidate as measured by its page rank in the DBpedia graph. An optimal model is trained using an SVM which combines these four features into an overall decision whether the named entity refers to a candidate resource or not.

@@

The input data for {\acronym} is a regular piece of text, for example a newspaper article in which entity-representing sequences of tokens (\emph{fragments}) are already identified (we use the Stanford NER~\cite{ner}). For a given fragment, the algorithm's task is to determine the matching real-world entity. It does so by picking the most likely entity from a set of entity candidates, e.g. specified in a knowledge base, in our case DBpedia.

@@

Candidates are scored by a support vector machine.\footnote{For the training set the correct assignment and three randomly chosen false candidates for each entity are used. {\acronym} is using Weka's SMO: \url{http://weka.sourceforge.net/doc.dev/weka/classifiers/functions/SMO.html}} Four features are computed for each candidate, which are used as input for the SVM: tf-idf vector similarity, semantic neighborhood signature vector similarity, candidate reference probability, and page rank. In the following we explain each of them in more detail.

% ------------ tf--idf vector similarity -------------------
\subsection{tf-idf vector similarity}

@@

To calculate the similarity score for each candidate, their semantic neighborhood signature vector needs to compared to the semantic neighborhood signature vector of the input. Unfortunately, the input does not have a semantic neighborhood signature of its own. As a surrogate, a candidate vector is generated as follows: Let $m(e) \rightarrow c$ be a mapping from an entity $e$ to a count $c$. For each candidate entity of every fragment, check whether it already appears in the mapping. If not, add it to the map and set its count to one. If it exists, increment its count by one.

@@

$$\mathrm{crp}(a, e) := \frac{\mathrm{ref}(a, e)}{{\sum\limits_{a'\in A} \mathrm{ref}(a', e)}}$$

Where $\mathrm{ref}(a, e)$ is the number of references from anchor/candidate $a$ to entity $e$ and $A$ is the set of all anchors referring to $e$.

@@

The idea behind the PageRank algorithm is to estimate the relevance of a website to a search query not only by whether the search terms appear on that page, but also by how many other pages were providing a hyperlink to the website in question. As DBpedia entries form a highly interconnected graph, the same principle can be applied in order to determine the page rank of a resource in DBpedia.

Let $u$ be a web page. $B_u$ is the set of pages that have a hyperlink pointing to $u$. Let $N_u$ be the number of outgoing links for page $u$. A slightly simplified PageRank is then defined as follows: 
$$R(u) = c \sum_{v \in B_u} \frac{R(v)}{N_v}$$
Where $c$ is a normalizing factor. Notably the PageRank of a page is derived from the PageRank of its incoming links, meaning that the algorithm has to be iterated until the system reaches a steady state. The overall PageRank for a graph is supposed to be constant, but there are corner cases (like cycles) that have to be accounted for; see for more information. Our implementation uses the JUNG framework\footnote{Java Universal Network/Graph Framework: \url{http://jung.sourceforge.net/}}.

@@

Table~\ref{table:redirects1} shows the evaluation results. 
\acronym{} outperforms both FOX and DBpedia Spotlight on the AIDA/CoNLL dataset and achieves comparable results on the KORE50 dataset. 
The latter is due to the fact that KORE50 was designed to have a high degree of ambiguity and short sentences, which renders {\acronym 's} approach less effective: Short sentences provide fewer tokens to evaluate a tf-idf overlap, while increased numbers of candidates allow for more detrimental semantic neighborhood signature vector overlaps.

@@

Table ~\ref{table:components} shows results leaving out single features and using only one feature. We clearly see that the page rank is the dominating factor. This deserves closer investigation.

@@

\begin{table}[t]
\begin{center}
\begin{tabularx}{\textwidth}{ X Y Y Y Y Y Y}
\toprule
& \multicolumn{3}{c}{AIDA/CoNLL}  
& \multicolumn{3}{c}{KORE50}\\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
 & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ 
\midrule
\emph{without} &  &  &  &  &  &  \\ 
SNS & 0.60 & 0.57 & 0.59 & 0.32 & 0.28 &  0.29 \\
tf-idf & 0.60 & 0.57 & 0.58 & 0.32 & 0.28 &  0.29 \\
page rank & 0.36 & 0.34 & 0.35 & 0.12 & 0.10 &  0.11 \\
CRP & 0.62 & 0.60 & 0.61 & 0.31 & 0.27 & 0.29 \\
\midrule
\emph{only with} &  &  &  &  &  &  \\ 
SNS & 0.36 & 0.35 & 0.36 & 0.07 & 0.06 &  0.06 \\
tf-idf & 0.22 & 0.21 & 0.22 & 0.08 & 0.07 &  0.07 \\
page rank & 0.62 & 0.60 & 0.61 & 0.32 & 0.28 &  0.29 \\
CRP & 0.21 & 0.20 & 0.21 & 0.08 & 0.07 &  0.07 \\
\bottomrule
\end{tabularx}
\end{center}
\caption{Results for the AIDA/CoNLL and KORE50 datasets when one of the components is removed and when only one component is used for classification. SNS is Semantic Neighborhood Signature Vector Similarity, CRP is Candidate Reference Probability.}
\label{table:components}
\end{table}

@@

We have presented {\acronym}, a new entity linking approach that exploits a combination of features that are combined using a Support Vector Machine. We have shown that the approach works very well when the input sentences are longer, clearly outperforming FOX and DBpedia Spotlight on the AIDA/CoNLL dataset. On the KORE50 dataset where sentences are shorter, it achieves comparable results to these approaches. In future work, we will release a service that can be publicly accessed using a similar interface to DBpedia Spotlight and FOX. We will also extend the approach to other languages and perform a more systematic evaluation using the GERBIL framework \cite{gerbil}.
